---
title: "Datathon"
author: "Zach Soriano"
date: "2025-04-04"
output:
  pdf_document: default
  html_document: default
---
# What is the most upvoted sentiment, topic, etc.?
# Getting the DF in order
```{r}
getwd()
reddits <- read.csv("reddit_data.csv")
class(reddits)
head(reddits)
library(dplyr)
library(ggplot2)


## take a random sample of 100 reddit posts, all with selftexts (for analysis)
new_reddit <- data.frame()
reddits_with_selftext <- reddits |> filter(!is.na(selftext) & selftext != "")
for(i in 1:100){
  new_reddit <- rbind(new_reddit, reddits_with_selftext[sample(nrow(reddits_with_selftext), 1, TRUE),])
}
head(new_reddit)
write.csv(new_reddit,"new_reddits.csv")

print(sentiment_vector_single_word)
length(sentiment_vector_single_word)
```

# what post classifications have the most upvotes? 
# adding classifications to DF
```{r}
for(i in 1:100){
  new_reddit$classification[i] = sentiment_vector_single_word[i]
}
# summarise mean and median upvotes for positive/negative posts
new_reddit |> group_by(classification) |> summarise(med_upvotes = median(ups), mean_upvotes = mean(ups))

# plot to show amount of outliers, for 'viral' posts
boxplot(new_reddit$ups ~ new_reddit$classification, xlab = "Classification", ylab = "Upvotes", ylim = c(0,10000))


```
Response: Out of the 100 Reddit posts sampled, positively classified posts are more likely to have more upvotes than negatively classified posts, with positively classified posts receiving mean upvote count of 3,979 upvotes and a median upvote count of 1,537 upvotes, compared to negatively classified posts which only received a mean and median upvote count of 2,578 and 1,408 respectively. Another factor which I would informally introduce as something I would call a "viral" factor, the likelihood that a posts receives enough upvotes to be considered an outlier. In this sample, negative posts are more likely to go viral, which is visible in the figure's outliers represented as dots outside of thr boxplot. 

# What is the relationship between comments and upvotes?
```{r}
# assignment
num_comments <- new_reddit$num_comments
num_ups <- new_reddit$ups

# create a linear regression, formula being upvotes predicted by 
# number of comments
line_reg_reddits <- lm(new_reddit$ups ~ new_reddit$num_comments)

# plot upvotes predicted by comments
plot(new_reddit$ups ~ new_reddit$num_comments, main = "Num of upvotes predicted by num of comments", ylab = "Number of upvotes", xlab = "Number of comments")
abline(line_reg_reddits, col = "red", lwd = 2)

# how much are these two variables correlated?
r <- cor(x = num_comments, y = num_ups)
r_squared <- r^2
r
r_squared
```
Response: From the 100 randomly sampled reddit posts, the relationship between the number of upvotes and the number of comments a post receives is positively correlated, meaning that generally the more comments a post receives the more likes it is likely to receive. However, these two variables share a correlation coefficient of just r = 0.404, and an r^2 value of just r^2 = .163, meaning that just 16% of the variation in the amount of upvotes a post receives is accounted for by the amount of comments a post receives. A general interpretation of this is that a post that receives a large amount of engagement in the comment section is more likely to have more upvotes, but because of the weak correlation between these two variables, we can assume that there are many more factors in which predict the amount of upvotes a post receives, rather than just looking at how many comments a post receives. 

# Do positive/negative posts get more engagement?
```{r}
# add engagement score to the DF, defined by number of likes+comments
df_with_engagement <- new_reddit |> mutate(engagement_score = ups + num_comments)

# mean and median engagement scores, for positive/negative posts
df_with_engagement |> group_by(classification) |> summarise(mean_engagement_score = mean(engagement_score), median_engagement_score = median(engagement_score))

# save the scores for plotting
score <- df_with_engagement$engagement_score
classification <- df_with_engagement$classification

# visualize the engagement score by positive/negative posts


```
Response: From the 100 randomly sampled reddit posts, positively classified posts boast much higher engagement scores than negatively classified posts. The mean and median scores for engagement posts that are positively classified are 4,199 and 2,1,555 respectively. Recall that engagement score is calculated with likes + comments. These greatly surpass the mean and median engagement scores of negatively classified posts, which were 2,888 and 1,452 respectively. A possible conclusion for this, if your goal was to make a post that receives a lot of upvotes and comments, is to make a post with positive content and messaged. 

# Predicting the type of post that will receive the most engagement
# using a classification (pos/neg) and cluster
```{r}
# Clusters to be defined as 
# 0: ["finance", "current", "state"],
# 1: ["politics", "concern", "disappointed"],
# 2: ["travel", "future", "concern"],
# 3: ["health", "sports", "downfall"],
# 4: ["education", "remains", "seen"],
# 5: ["technology", "disappointed", "future"],
# 6: ["health", "new", "love"],
# 7: ["environment", "disappointed", "remains"]


# Cluster assignment, with the code provided in the collab. Each
# element was fed into the assignment algorithm, and each element of
# 'cluster_assignment' corresponds to the cluster of the same element in 'new_reddit'.

cluster_assignment <- c(
  6, 0, 5, 4, 2, 3, 7, 1, 0, 7, 1, 6, 4, 5, 2, 5, 6, 2, 6, 0,
  3, 0, 7, 5, 5, 6, 1, 1, 6, 5, 5, 0, 3, 6, 6, 2, 0, 6, 1, 0,
  6, 3, 4, 5, 2, 6, 6, 5, 3, 2, 0, 2, 4, 1, 2, 5, 1, 7, 7, 4,
  0, 1, 1, 1, 5, 3, 0, 4, 2, 1, 0, 4, 2, 3, 0, 7, 4, 7, 4, 7,
  7, 7, 3, 2, 1, 2, 4, 4, 5, 7, 3, 4, 3, 3, 7, 3, 3, 3, 4, 1
)
# add the cluster to the DF
for(i in 1:100){
  new_reddit$cluster[i] = cluster_assignment[i]
}

# from before, adding engagement score to the DF
ups <- new_reddit$ups
comments <- new_reddit$num_comments
for(i in 1:100){
  new_reddit$engagement_score[i] = comments[i] + ups[i]
}

# mean and median engagement scores for each cluster
new_reddit |> group_by(cluster) |> summarise(mean_engagement_score = mean(engagement_score), median_engagement_score = median(engagement_score))

# visualizing the engagement score by cluster
boxplot(new_reddit$engagement_score ~ new_reddit$cluster, 
        xlab = "Cluster", ylab = "Engagement Score", 
        main = "Engagement Score by Cluster", ylim = c(0,10000))

```
Response: In order to very broadly predict which posts will have good engagement, we consider a few assumptions. The first assumption is that the algorithm correctly assigned the post to a cluster based on keywords in the post. Another is that the conclusion we made before about positively classified posts being more successful in terms of engagement compared to negative posts. Lastly we assume that the 100 randomly chosen posts are representative of posts with similar content in the entire dataset of reddit posts. Under all of these assumptions, we consider the following conclusion about which posts will have the most engagement. The clusters with the highest median engagement are clusters 0, 2, 4, and 7. Note that we consider the median engagement score because one post that receives a very unlikely amount of engagement (like the post in cluster 7 with almost 50,000 engagement score) will ruin the validity of the mean. Narrowing down the clusters to just 4, we can consider all of the topics in each of these clusters. The 12 topics are finance, current, state, travel, future, concern, education, remains, seen, environment, disappointed, and remains. Recall from earlier we concluded that posts with positive classifications are more likely to receive better engagement scores. Thus, if your goal was to make a post with the most engagement, we would conclude from our findings that a post that is overall written with a positive tone about one of the 12 topics listed above would be your best bet. Notice how some of the topics are inherently negative, so the classification of your post is likely correlated with your topic. For example, the successful negatively classified posts are likely about the negatively associated topics, like disappointed, or concern. All other "successful" topics would likely be written with a positive classification. 

# How does engagement differ across subreddits? Hypothesis: A post posted in a subreddit with more subscribers will have a higher engagement score
```{r}
# Sort each post by its engagement, which will be compared to the number of subscribers in the post's subreddit
new_reddit |> arrange(engagement_score) |> select(engagement_score, subreddit_subscribers)
# from this, it is not obvious to say that subreddit subscribers are # the likely cause of a higher engagement scores

# now to see the relationship visually
lm_subreddits <- lm(new_reddit$engagement_score ~ new_reddit$subreddit_subscribers)
plot(new_reddit$engagement_score ~ new_reddit$subreddit_subscribers, main = "Engagement by Subreddit Subscribers", xlab = "Number of Subreddit Subscribers", ylab = "Engagement Score")
abline(lm_subreddits, col = "red", lwd = 2)

cor(y = new_reddit$engagement_score, x = new_reddit$subreddit_subscribers)

# due to an low correlation coefficient, we observe interestingly enough that the the size of the subreddit has little to no observable effect on the engagement of the post. 

```



